---
title: "analysis"
author: "KSA"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  error = FALSE,
  fig.width = 12,
  fig.height = 10,
  cache = FALSE
)
require("patchwork")
require("data.table")
require("ampvis2")
require("magrittr")
require("stringr")

#knitting will always be done in the context of the Rmd file, otherwise not
if (interactive()) {
  if (!grepl("manuscript_analysis$", getwd())) {
    setwd("manuscript_analysis")
  }
}
source("functions.R")
```

# Prediction accuracy across WWTPs

## top 100 ASVs, 10 iterations, 200 epochs, smoothing factor 8
```{r run_20220420}
plot_all("results/20220420")
# "only_pos_func": false,
# "pseudo_zero": 0.01,
# "max_zeros_pct": 0.60,
# "top_n_taxa": 100,
# "num_features": 10,
# "iterations": 10,
# "max_epochs_lstm": 200,
# "window_size": 10,
# "num_clusters_idec": 10,
# "tolerance_idec": 0.001,
# "splits": [
#     0.75,
#     0.10,
#     0.15
# ]
```

## 1 iteration, 1000 max epochs
```{r run_20220421}
# "only_pos_func": false,
# "pseudo_zero": 0.01,
# "max_zeros_pct": 0.60,
# "top_n_taxa": 100,
# "num_features": 10,
# "iterations": 1,
# "max_epochs_lstm": 1000,
# "window_size": 10,
# "num_clusters_idec": 10,
# "tolerance_idec": 0.001,
# "splits": [
#     0.75,
#     0.10,
#     0.15
# ]
plot_all("results/20220421")
```

## 10 iterations, 2000 max epochs, window size 20
```{r run_20220422}
# "only_pos_func": false,
# "pseudo_zero": 0.01,
# "max_zeros_pct": 0.60,
# "top_n_taxa": 100,
# "num_features": 10,
# "iterations": 10,
# "max_epochs_lstm": 2000,
# "window_size": 20,
# "num_clusters_idec": 10,
# "tolerance_idec": 0.001,
# "splits": [
#     0.75,
#     0.10,
#     0.15
# ]
plot_all("results/20220422")
```

## top 200 ASVs, windows size 10, 20 IDEC clusters
```{r run_20220427}
# "only_pos_func": false,
# "pseudo_zero": 0.01,
# "max_zeros_pct": 0.60,
# "top_n_taxa": 200,
# "num_features": 10,
# "iterations": 10,
# "max_epochs_lstm": 2000,
# "window_size": 10,
# "num_clusters_idec": 20,
# "tolerance_idec": 0.001,
# "splits": [
#     0.75,
#     0.10,
#     0.15
# ]
plot_all("results/20220427")
```

## 5 IDEC clusters
```{r run_20220429}
# "only_pos_func": false,
# "pseudo_zero": 0.01,
# "max_zeros_pct": 0.60,
# "top_n_taxa": 200,
# "num_features": 10,
# "iterations": 10,
# "max_epochs_lstm": 2000,
# "window_size": 10,
# "num_clusters_idec": 5,
# "tolerance_idec": 0.001,
# "splits": [
#     0.75,
#     0.10,
#     0.15
# ]
plot_all("results/20220429")
```

# smoothing factor 4
```{r run_20220506}
# "only_pos_func": false,
# "pseudo_zero": 0.01,
# "max_zeros_pct": 0.60,
# "top_n_taxa": 200,
# "num_features": 10,
# "iterations": 10,
# "max_epochs_lstm": 2000,
# "window_size": 10,
# "num_clusters_idec": 5,
# "tolerance_idec": 0.001,
# "smoothing_factor": 4,
# "splits": [
#     0.75,
#     0.10,
#     0.15
# ]
plot_all("results/20220506")
```

```{r updated_data}
# "metadata_date_col": "Date",
# "tax_level": "OTU",
# "tax_add": ["Species", "Genus"],
# "functions": [
#     "AOB",
#     "NOB",
#     "PAO",
#     "GAO",
#     "Filamentous"
# ],
# "only_pos_func": false,
# "pseudo_zero": 0.01,
# "max_zeros_pct": 0.60,
# "top_n_taxa": 200,
# "num_features": 10,
# "iterations": 10,
# "max_epochs_lstm": 2000,
# "window_size": 10,
# "num_clusters_idec": 5,
# "tolerance_idec": 0.001,
# "smoothing_factor": 4,
# "splits": [
#     0.75,
#     0.10,
#     0.15
# ]
plot_all("results/20220511_updateddata")
```


```{r updated_data_novalidation}
# "functions": [
#     "AOB",
#     "NOB",
#     "PAO",
#     "GAO",
#     "Filamentous"
# ],
# "only_pos_func": false,
# "pseudo_zero": 0.01,
# "max_zeros_pct": 0.60,
# "top_n_taxa": 200,
# "num_features": 10,
# "iterations": 10,
# "max_epochs_lstm": 2000,
# "window_size": 10,
# "num_clusters_idec": 5,
# "tolerance_idec": 0.001,
# "smoothing_factor": 4,
# "splits": [
#     0.85,
#     0.0,
#     0.15
# ]
plot_all("results/20220518_updateddata")
```

# colored
```{r PCOA_colored}
#data set exactly same settings as 20220506, just with additional data output
results_dir <- "results/20220510/results_20220510_190511"
AAW_20220510 <- combine_abund(
  results_dir,
  cluster_type = "abund"
)

AAW_20220510_reformatted <- load_data_reformatted(results_dir)

# run data (here smoothing factor 4)
amp_ordinate(
  AAW_20220510,
  type = "pcoa",
  distmeasure = "bray",
  sample_color_by = "split_dataset",
  sample_trajectory = "Date"
) +
  scale_color_manual(
    values = c("grey50", RColorBrewer::brewer.pal(6, "Paired")[c(4:6)])
  ) +
  theme(legend.title = element_blank())
```

# (updated data) Aalborg West comparison of true vs predicted (smoothing factor 4)
## Principal Coordinates Analysis
```{r updated_CA_AAW_4smooth}
# Configuration:
# {
#     "abund_file": "data/datasets/Aalborg W/ASVtable.csv",
#     "taxonomy_file": "data/datasets/Aalborg W/taxonomy.csv",
#     "metadata_file": "data/metadata.csv",
#     "results_dir": "results",
#     "metadata_date_col": "Date",
#     "tax_level": "OTU",
#     "tax_add": ["Species", "Genus"],
#     "functions": [
#         "AOB",
#         "NOB",
#         "PAO",
#         "GAO",
#         "Filamentous"
#     ],
#     "only_pos_func": false,
#     "pseudo_zero": 0.01,
#     "max_zeros_pct": 0.60,
#     "top_n_taxa": 200,
#     "num_features": 10,
#     "iterations": 10,
#     "max_epochs_lstm": 2000,
#     "window_size": 10,
#     "num_clusters_idec": 5,
#     "tolerance_idec": 0.001,
#     "smoothing_factor": 4,
#     "splits": [
#         0.85,
#         0.0,
#         0.15
#     ]
# }
results_batch_dir <- "results/20220518_updateddata"
results_dir <- file.path(results_batch_dir, "results_20220513_200906")
AAW_20220518 <- combine_abund(
  results_dir,
  cluster_type = "abund"
)

AAW_20220518_reformatted <- load_data_reformatted(results_dir)


amp_ordinate(
  AAW_20220518,
  type = "pcoa",
  distmeasure = "bray",
  sample_color_by = "split_dataset",
  sample_trajectory = "Date"
) +
  scale_color_manual(
    values = c("grey50", RColorBrewer::brewer.pal(6, "Paired")[c(4, 6)])
  ) +
  theme(legend.title = element_blank())
```

## (updated data) Time Series example ASV1
```{r updated_TS_AAW_4smooth}
AAW_20220518_long <- amp_export_long(
  AAW_20220518,
  tax_levels = "OTU",
  metadata_vars = c("Date", "split_dataset")
)

ASV1_timeseries <- ggplot(
  AAW_20220518_long[grepl("^ASV1;", OTU)],
  aes(
    x = Date,
    y = count,
    color = split_dataset
  )
) +
  geom_line() +
  scale_color_manual(
    values = c("grey10", RColorBrewer::brewer.pal(6, "Paired")[c(4, 6)])
  )

ASV2_timeseries <- ggplot(
  AAW_20220518_long[grepl("^ASV2;", OTU)],
  aes(
    x = Date,
    y = count,
    color = split_dataset
  )
) +
  geom_line() +
  scale_color_manual(
    values = c("grey10", RColorBrewer::brewer.pal(6, "Paired")[c(4, 6)])
  ) +
  scale_x_date(
    date_breaks = "3 months",
    date_labels =  "%b %Y"
  ) +
  theme(
    axis.title = element_blank(),
    axis.text.x = element_text(angle = 90, hjust = 1)
  )

ggsave(ASV1_timeseries, file = "ASV1_timeseries.png", width = 20, height = 8)
ggsave(ASV2_timeseries, file = "ASV2_timeseries.png", width = 20, height = 8)


######################################
rects <- data.frame(xstart = seq(0,80,20), xend = seq(20,100,20), color = letters[1:5])


#As Baptiste points out, the order of the geom's matters, so putting your data as last will 
#make sure that it is plotted "on top" of the background rectangles. Updated code, but
#did not update the JPEG...I think you'll get the point.

ggplot() + 
  geom_rect(data = rects, aes(xmin = xstart, xmax = xend, ymin = -Inf, ymax = Inf, fill = color), alpha = 0.4)
#############################
```

```{r fivenum_BC_ASV}
runs <- list.files(
  results_batch_dir,
  pattern = "^results_.*",
  full.names = TRUE
)

d_list <- lapply(runs, read_results)
names(d_list) <- runs
combined <- rbindlist(
  d_list,
  idcol = "results_folder",
  fill = TRUE
)[
  !is.na(cluster_type) & value > 0
]
combined[cluster_type == "Single ASV" & error_metric == "Bray Curtis", fivenum(value)]
```

```{r fivenum_reads}
#five number statistics of sum of reads per data set
list.dirs(
  "results/20220518_updateddata",
  full.names = TRUE,
  recursive = FALSE
) %>%
  lapply(function(dataset) {
    abund <- fread(
      file.path(dataset, "data_reformatted", "abundances.csv"),
      drop = 1
    )
    fivenum(rowSums(abund))
  })
```
