---
title: "analysis"
author: "KSA"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  error = FALSE,
  fig.width = 12,
  fig.height = 10,
  cache = FALSE
)
require("patchwork")
require("data.table")
require("ampvis2")
require("magrittr")
require("stringr")
require("lubridate")
require("tidygraph")
require("ggraph")
require("igraph")
require("patchwork")
require("scales")

#when knitting will always be done in the context of the Rmd file, otherwise not
if (interactive()) {
  if (!grepl("analysis$", getwd())) {
    setwd("analysis")
  }
} else {
  renv::load("..")
}

source("functions.R")
set.seed(42)
# main batch run using run_loopdatasets.bash or run_predwindows.bash
# wrapper scripts. Prediction windows tests are hard coded
results_batch_dir <- "results/allwwtps"

# cluster type to use for examples
cluster_type <- "graph"
```

# five number statistics of sum of reads per data set
```{r fivenum_reads}
datasets <- list.dirs(
  results_batch_dir,
  full.names = TRUE,
  recursive = FALSE
)
readstats <- datasets %>%
  lapply(function(dataset) {
    abund <- fread(
      file.path(dataset, "data_reformatted", "abundances.csv"),
      drop = 1
    )
    fivenum(rowSums(abund))
  })
names(readstats) <- basename(datasets)
as.data.frame(readstats, check.names = FALSE)
```

# Prediction accuracy across WWTPs
## Bray-Curtis only
```{r}
bp_all_list <- boxplot_all(
  results_batch_dir,
  save = TRUE
)


## Bray-Curtis only, flip x and y (otherwise too big plot for manuscript)
runs <- list.dirs(
  results_batch_dir,
  full.names = TRUE,
  recursive = FALSE
)

d_list <- lapply(
  runs,
  read_performance,
  add_dataset_info = "numsamples"
)
names(d_list) <- runs
combined <- rbindlist(
  d_list,
  idcol = "results_folder",
  fill = TRUE
)[
  !is.na(cluster_type) & value > 0
]

# copied from boxplot_all() just for this plot
#order datasets by the number of samples (extracted from dataset names)
combined[,unique(dataset)] %>%
  stringi::stri_extract_all_regex("\\(.*$", "") %>%
  stringi::stri_extract_all_regex("[0-9]+") %>%
  as.numeric -> nsamples
datasets_ordered <- combined[
  ,
  unique(dataset)
][
  order(nsamples, decreasing = F)
]
combined[, dataset := factor(dataset, levels = datasets_ordered)]
bp_all_BC <- ggplot(
  combined[error_metric == "Bray Curtis"],
  aes(
    value,
    cluster_type,
    color = cluster_type
  )
) +
  geom_boxplot(
    outlier.shape = 1,
    outlier.size = 1
  ) +
  facet_grid(
    rows = vars(dataset),
    scales = "free"
  ) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = 14),
    axis.text.y = element_blank(),
    axis.text.x = element_text(size = 14),
    axis.title = element_blank(),
    axis.ticks.y = element_blank(),
    #strip.text.x = element_blank(),
    strip.text.y = element_text(size = 14, angle = 0),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_line(color = "gray70"),
    panel.background = element_rect(fill = "white")
  ) +
  scale_x_continuous(
    trans = "sqrt",
    breaks = c(0.0, 0.05, 0.1, 0.15, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8)
  ) +
  scale_color_brewer(palette = "Set2")

ggsave(
  file.path(results_batch_dir, "boxplot_all_BC.png"),
  plot = bp_all_BC,
  width = 10,
  height = 15
)
```

## BC+MAE+MSE
```{r}
bp_all_list <- boxplot_all(
  results_batch_dir,
  save = FALSE
)

# customize axis breaks a bit for the specific data
bp_all_list[["Bray Curtis"]] <- bp_all_list[["Bray Curtis"]] +
  scale_y_continuous(
    trans = "sqrt",
    breaks = c(0.5, 0.1, 0.15, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8)
  )
bp_all_list[["Mean Absolute Error"]] <- bp_all_list[["Mean Absolute Error"]] +
  scale_y_continuous(
    trans = "sqrt",
    breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1.0, 2, 3, 4, 5, 6)
  )
bp_all_list[["Mean Squared Error"]] <- bp_all_list[["Mean Squared Error"]] +
  scale_y_continuous(
    trans = "sqrt",
    breaks = c(0.5, 1, 5, 10, 20, 30, 40)
  )

# Compose all plots in the list using patchwork
bp_all_plot <- purrr::reduce(
  bp_all_list,
  `/`
)
ggsave(
  file.path(results_batch_dir, "boxplot_all.png"),
  plot = bp_all_plot,
  width = 16,
  height = 12
)
```


## Check some clusters from AAE
```{r}
plot_performance(
  file.path(results_batch_dir, "Aalborg E")
)
```

# Prediction window test (note: not using results_batch_dir variable)
## Aalborg East
```{r}
boxplot_all(
  "results/predwindow_aae",
  add_dataset_info = "predwindow"
)
```

## Randers
```{r}
boxplot_all(
  "results/predwindow_randers",
  add_dataset_info = "predwindow",
  plot_width = 8,
  plot_height = 8
)
```

## Mariagerfjord
```{r}
bp_mar_predwin_list <- boxplot_all(
  "results/predwindow_mariagerfjord",
  add_dataset_info = "predwindow",
  save = FALSE
)


bp_mar_predwin_list[["Bray Curtis"]] <- bp_mar_predwin_list[["Bray Curtis"]] +
  scale_y_continuous(
    trans = "sqrt",
    breaks = c(0.0, 0.05, 0.1, 0.15, 0.2, 0.3, 0.5, 0.7)
  ) + expand_limits(y = 0.0)
bp_mar_predwin_list[["Mean Absolute Error"]] <- bp_mar_predwin_list[["Mean Absolute Error"]] +
  scale_y_continuous(
    trans = "sqrt",
    breaks = c(0.0, 0.5, 1.0, 2.0, 3.0, 4.0)
  ) + expand_limits(y = 0.0)
bp_mar_predwin_list[["Mean Squared Error"]] <- bp_mar_predwin_list[["Mean Squared Error"]] +
  scale_y_continuous(
    trans = "sqrt",
    breaks = c(0, 1, 5, 10, 20, 30, 50, 80)
  ) + expand_limits(y = 0.0)
bp_mar_predwin_plot <- purrr::reduce(
  bp_mar_predwin_list,
  `/`
)
ggsave(
  "results/predwindow_mariagerfjord/boxplot_all.png",
  plot = bp_mar_predwin_plot,
  width = 8,
  height = 10
)
```

## Timeseries examples (predwindow)
```{r}
dir <- "results/predwindow_mariagerfjord"
runs <- list.dirs(
  dir,
  recursive = FALSE,
  full.names = TRUE
)
predwindowtest <- lapply(
  runs,
  function(rundir) {
    d <- amp_export_long(
      combine_abund(
        rundir,
        cluster_type = cluster_type
      ),
      tax_levels = "OTU",
      metadata_vars = c("Date", "split_dataset")
    )
    d[, dataset := basename(rundir)]
  }
)
d <- rbindlist(predwindowtest)
d[, dataset := gsub("^[^0-9]* +", "", dataset)]
d[, dataset := factor(
  dataset,
  levels = unique(dataset)[order(as.integer(gsub("[^0-9]*", "", unique(dataset))))])
]

#ASV1
ASV1 <- d[grepl("^ASV1;", OTU)]
plot_asv1 <- plot_timeseries(
  ASV1,
  save = FALSE
) +
  facet_grid(rows = vars(dataset))
ggsave(
  file = file.path(
    dir,
    "predwindow_mariagerfjord_asv1.png"
  ),
  plot = plot_asv1
)

#ASV2
ASV2 <- d[grepl("^ASV2;", OTU)]
plot_asv2 <- plot_timeseries(
  ASV2,
  save = FALSE
) +
  facet_grid(rows = vars(dataset))
ggsave(
  file = file.path(
    dir,
    "predwindow_mariagerfjord_asv2.png"
  ),
  plot = plot_asv2
)

#ASV24
ASV24 <- d[grepl("^ASV24;", OTU)]
plot_asv24 <- plot_timeseries(
  ASV24,
  save = FALSE
) +
  facet_grid(rows = vars(dataset))
ggsave(
  file = file.path(
    dir,
    "predwindow_mariagerfjord_asv24.png"
  ),
  plot = plot_asv24
)


#plot_asv24 <- plot_timeseries(ASV24, save = FALSE) +
#  theme(legend.position = "bottom") +
#  ggtitle("ASV24 - Nitrospira, defluvii")

```

# PCoA (all WWTPs loop)
```{r PCOA_colored_all}
wwtp_results_dirs <- list.dirs(
  results_batch_dir,
  full.names = TRUE,
  recursive = FALSE
)

plots <- lapply(
  wwtp_results_dirs,
  function(wwtp) {
    plot <- wwtp %>%
      combine_abund(cluster_type = cluster_type) %>%
      amp_ordinate(
        type = "pcoa",
        distmeasure = "bray",
        transform = "none",
        filter_species = 0,
        sample_color_by = "split_dataset",
        sample_trajectory = "Date",
        sample_trajectory_group = "Date"
      ) +
      scale_color_manual(
        values = c("grey80", RColorBrewer::brewer.pal(3, "Set2")[c(2,3,1)]),
        labels = c(
          real = "Real",
          train = "Prediction - Train",
          val = "Prediction - Validation",
          test = "Prediction - Test"
        ),
        breaks = c(
          "real",
          "train",
          "val",
          "test"
        )
      ) +
      scale_x_continuous(breaks = breaks_pretty(n = 3)) +
      scale_y_continuous(breaks = breaks_pretty(n = 3)) +
      theme(
        legend.title = element_blank(),
        legend.text = element_text(size = 16),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 16),
        legend.position = "bottom"
      ) +
      coord_fixed(ratio = 1)

    # real_data <- data.table(plot$data)[predicted == "real"]
    # # extract first and last row (corresponds to first and last sample in time)
    # labels_data <- rbindlist(list(real_data[1, ], real_data[.N, ]))
    # plot <- plot + geom_text(aes(label = Date, color = "black"), data = labels_data)

    ggsave(
      filename = file.path(
        results_batch_dir,
        paste0("PCoA - ", basename(wwtp), ".png")
      ),
      plot = plot,
      width = 10,
      height = 10
    )
    return(plot)
  }
)
names(plots) <- basename(wwtp_results)
#hideme <- lapply(plots, print)
```

## All WWTPs test data (real+pred) combined
```{r}
wwtp_results_dirs <- list.dirs(
  results_batch_dir,
  full.names = TRUE,
  recursive = FALSE
)#[8] # Egå sometimes has multiple samples per week

#list of all data (train+val+test) from all WWTPs
allwwtps_testcompare_list <- lapply(
  wwtp_results_dirs,
  function(results_dir) {
    d <- combine_abund(
      results_dir,
      cluster_type = cluster_type
    )
    test_predicted <- suppressMessages({
      amp_filter_samples(
        d,
        split_dataset %in% "test",
        predicted %in% "predicted"
      )
    })
    test_real <- suppressMessages({
      amp_filter_samples(
        d,
        split_dataset %in% "real",
        Date %in% test_predicted$metadata$Date
      )
    })
    test_combined <- amp_merge_ampvis2(
      test_predicted,
      test_real,
      by_refseq = FALSE
    )
    test_combined$metadata$predicted <- factor(
      test_combined$metadata$predicted,
      levels = c("real", "predicted")
    )
    test_combined_long <- amp_export_long(
      test_combined,
      tax_levels = "OTU",
      metadata_vars = c("Date", "predicted")
    )

    # if/when there are multiple samples per time point, use the mean, and go cast straight away
    test_combined_long <- dcast(
      test_combined_long[, .(count = mean(count)), by = c("predicted", "OTU", "Date")],
      Date+OTU ~ predicted,
      value.var = "count"
    )
    OTU_functions <- fread(
      file.path(results_dir, "data_reformatted", "taxonomy_wfunctions.csv"),
      select = c(1, 9:13),
      data.table = FALSE
    )
    test_combined_long_func <- merge(
      test_combined_long,
      OTU_functions,
      by = "OTU",
      all.x = TRUE,
      all.y = FALSE
    )
    return(test_combined_long_func)
  }
)
names(allwwtps_testcompare_list) <- basename(wwtp_results_dirs)
allwwtps_testcompare <- rbindlist(allwwtps_testcompare_list, idcol = "dataset")
allwwtps_testcompare <- tidyr::separate(allwwtps_testcompare, col = "OTU", into = c("OTU", "Species", "Genus"), sep = "; ")
setDT(allwwtps_testcompare)
allwwtps_testcompare[AOB == "POS", genus_function := "AOB"]
allwwtps_testcompare[NOB == "POS", genus_function := "NOB"]
allwwtps_testcompare[PAO == "POS", genus_function := "PAO"]
allwwtps_testcompare[GAO == "POS", genus_function := "GAO"]
allwwtps_testcompare[Filamentous == "POS", genus_function := "Filamentous"]
allwwtps_testcompare[, dataset := as.factor(dataset)]
allwwtps_testcompare[, genus_function := as.factor(genus_function)]
allwwtps_testcompare <- allwwtps_testcompare[real > 0]
allwwtps_testcompare[predicted == 0, predicted := 0.001]
allwwtps_testcompare[
  !is.na(genus_function),
  MSE := mean((real - predicted)^2),
  by = c("dataset", "genus_function", "OTU")
]
allwwtps_testcompare[
  !is.na(genus_function),
  MAE := mean(abs(real - predicted)),
  by = c("dataset", "genus_function", "OTU")
]
allwwtps_testcompare[
  !is.na(genus_function),
  AE := abs(real - predicted),
  by = c("dataset", "genus_function", "OTU")
]
allwwtps_testcompare[
  !is.na(genus_function),
  MAPE := mean(abs((real - predicted)) / real) * 100,
  by = c("dataset", "genus_function", "OTU")
]

# summary of real/predicted errors, all datasets
allwwtps_testcompare[
  !is.na(genus_function),
  .(
    medianMSE = median(MSE, na.rm = TRUE),
    medianMAE = median(MAE, na.rm = TRUE),
    medianAE = median(AE, na.rm = TRUE),
    medianMAPE = median(MAPE, na.rm = TRUE)
  ),
  by = c("genus_function", "dataset")
]

# summary of real/predicted errors, genus functions only
allwwtps_testcompare[
  !is.na(genus_function),
  .(
    medianMSE = median(MSE, na.rm = TRUE),
    medianMAE = median(MAE, na.rm = TRUE),
    medianAE = median(AE, na.rm = TRUE),
    medianMAPE = median(MAPE, na.rm = TRUE)
  ),
  by = c("genus_function")
]

# summary of taxonomy
allwwtps_testcompare[
  ,
  .(
    uniqueOTU = uniqueN(OTU, na.rm = TRUE),
    uniqueGenus = uniqueN(Genus, na.rm = TRUE),
    uniqueSpecies = uniqueN(Species, na.rm = TRUE)
  ),
  by = "dataset"
]

# number of unique ASVs with a known function across ALL datasets
allwwtps_testcompare[!is.na(genus_function), uniqueN(OTU)]

# total unique ASVs  across ALL datasets
allwwtps_testcompare[, uniqueN(OTU)]

# plotty plot
plot_genus_func <- ggplot(
  allwwtps_testcompare[!is.na(genus_function)],
  aes(x = MAPE, y = dataset, group = dataset)
) +
  geom_boxplot() +
  scale_x_log10(breaks = c(5, 10, 20, 50, 100, 300)) +
  facet_grid(cols = vars(genus_function)) +
  theme(axis.title.y = element_blank()) +
  theme(panel.grid.minor = element_blank())
plot_genus_func

ggsave(
  plot = plot_genus_func,
  filename = file.path(
    results_batch_dir,
    "functions_acc.png"
  ),
  width = 12,
  height = 10
)

ggplot(
  allwwtps_testcompare[!is.na(genus_function)],
  aes(x = MAPE, y = genus_function, group = genus_function)
) +
  geom_boxplot() +
  scale_x_log10(breaks = c(5, 10, 20, 50, 100, 300)) +
  theme(axis.title.y = element_blank()) +
  theme(panel.grid.minor = element_blank())
plot_genus_func
```

# Time Series examples
```{r timeseries}
wwtp_results_dirs <- list.dirs(
  results_batch_dir,
  full.names = TRUE,
  recursive = FALSE
)

#list of all data (train+val+test) from all WWTPs
plots <- lapply(
  wwtp_results_dirs,
  function(wwtp) {
    combined <- combine_abund(
      wwtp,
      cluster_type = cluster_type
    )
    long <- amp_export_long(
      combined,
      tax_levels = "OTU",
      metadata_vars = c("Date", "split_dataset")
    )
    
    #check normal distribution
    #qqplot(ASV1, rnorm(1000, mean = mean(ASV1), sd = sd(ASV1)))
    #qqplot(ASV2, rnorm(1000, mean = mean(ASV2), sd = sd(ASV2)))
    ASV1 <- long[grepl("^ASV1;", OTU)]
    if(nrow(ASV1) > 0) {
      plot_asv1 <- plot_timeseries(ASV1, save = FALSE) +
        theme(
          legend.position = "none",
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.y = element_blank()
        ) +
        ggtitle("ASV1 - Tetrasphaera, midas_s_5") +
        scale_y_continuous(breaks = breaks_pretty(n = 5))
    } else {
      plot_asv1 <- plot_spacer()
    }

    ASV2 <- long[grepl("^ASV2;", OTU)]
    if(nrow(ASV2) > 0) {
      plot_asv2 <- plot_timeseries(ASV2, save = FALSE) +
        theme(
          legend.position = "none",
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank()
        ) +
        ggtitle("ASV2 - Ca. Microthrix, parvicella") +
        scale_y_continuous(breaks = breaks_pretty(n = 5))
    } else {
      plot_asv2 <- plot_spacer()
    }

    ASV24 <- long[grepl("^ASV24;", OTU)]
    if(nrow(ASV24) > 0) {
      plot_asv24 <- plot_timeseries(ASV24, save = FALSE) +
        theme(
          legend.position = "bottom",
          legend.text = element_text(size = 16),
          axis.title.y = element_blank()
        ) +
        ggtitle("ASV24 - Nitrospira, defluvii") +
        scale_y_continuous(breaks = breaks_pretty(n = 5))
    } else {
      plot_asv24 <- plot_spacer()
    }

    prediction_example <- plot_asv1 / plot_asv2 / plot_asv24
    ggsave(
      plot = prediction_example,
      filename = file.path(
        results_batch_dir,
        paste0("timeseries example - ", basename(wwtp), ".png")
      ),
      width = 10,
      height = 12
    )

    #plot_timeseries(ASV1, save = TRUE)
    #plot_timeseries(ASV2, save = TRUE)
    #plot_timeseries(ASV3, save = TRUE)
    return(prediction_example)
  }
)

```

# Heatmap of real vs predicted
```{r}
results_dir <- file.path(results_batch_dir, "Mariagerfjord")
d <- combine_abund(
  results_dir,
  cluster_type = cluster_type
)
test_predicted <- amp_filter_samples(
  d,
  split_dataset %in% "test",
  predicted %in% "predicted"
)
test_real <- amp_filter_samples(
  d,
  split_dataset %in% "real",
  Date %in% test_predicted$metadata$Date
)
test_combined <- amp_merge_ampvis2(
  test_predicted,
  test_real,
  by_refseq = FALSE
)
test_combined$metadata$predicted <- factor(
  test_combined$metadata$predicted,
  levels = c("real", "predicted")
)

# manually clean up long taxonomy strings and reverse the order
# this is NOT generic for other datasets, it's just for the manuscript
clean_tax <- function(charfctr) {
  charfctr <- gsub(
    "s__|g__|Ca_[^_]+_|Sulfuritalea_",
    "",
    charfctr
  )
  charfctr <- sapply(
    strsplit(
      charfctr,
      "; "
    ),
    function(x)
      paste0(rev(x), collapse = "; "),
    simplify = TRUE
  )
  return(charfctr)
}
test_combined$tax$OTU <- clean_tax(test_combined$tax$OTU)
rownames(test_combined$tax) <- clean_tax(rownames(test_combined$tax))
rownames(test_combined$abund) <- clean_tax(rownames(test_combined$abund))

realvspredictedheatmap <- amp_heatmap(
  test_combined,
  tax_aggregate = "OTU",
  group_by = c("predicted"),
  normalise = FALSE,
  tax_show = 30,
  plot_values = FALSE,
  facet_by = "Date"
) +
  theme(
    strip.text = element_text(angle = 90, size = 14),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 16),
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.position = "bottom"
  ) +
    labs(fill = "% Relative Abundance") +
    scale_fill_gradientn(
      colours = rev(RColorBrewer::brewer.pal(3, "RdYlBu")),
      trans = "sqrt",
      na.value = "#67A9CF",
      oob = scales::squish,
      limits = c(0, 5),
      breaks = c(0, 1, 2, 3, 5),
      labels = c(0, 1, 2, 3, 5)
    ) +
    guides(fill = guide_colourbar(ticks.colour = "black"))

ggsave(
  plot = realvspredictedheatmap,
  filename = file.path(
    results_dir,
    "..",
    paste0("realvspredictedheatmap_", basename(results_dir), ".png")
  ),
  width = 20,
  height = 12
)
```

# Network plots
```{r}
graph_matrix_dir <- file.path(results_batch_dir, "Mariagerfjord/graph_matrix")
graph_matrix_list <- list.files(
  graph_matrix_dir,
  pattern = "graph_cluster_*",
  full.names = TRUE
)
#list is sorted lexicographically, reorder numerically
order <- graph_matrix_list %>%
  basename %>%
  stringi::stri_extract_all_regex("[0-9]+") %>%
  unlist(use.names = FALSE) %>%
  as.integer %>%
  order
graph_matrix_list <- graph_matrix_list[order]

plot_list <- lapply(
  graph_matrix_list[1:4],
  plot_graph
)

#plots are named after the filenames
plot <- wrap_plots(
  plot_list,
  guides = "collect",
  tag_level = "keep"
)
plot
ggsave(
  plot = plot,
  filename = file.path(
    graph_matrix_dir,
    "..",
    "graph_network_plot_mariagerfjord.png"
  )
)
```

# Statistical tests of real vs predicted
```{r}
wwtp_results <- list.dirs(
  results_batch_dir,
  full.names = TRUE,
  recursive = FALSE
)
statistics <- lapply(
  wwtp_results,
  function(wwtp) {
    cli::cat_print(wwtp)
    #read predicted data (incl train+val)
    combined <- combine_abund(
      wwtp,
      cluster_type = cluster_type
    )
    combined
    if(is.null(combined))
      return(NULL)
    #extract predicted samples (test set only)
    predicted <- suppressMessages({
      amp_filter_samples(
      combined,
      split_dataset %in% "test",
      predicted %in% "predicted"
    )
    })
    #extract real samples from the same dates as the predicted
    dates <- predicted[["metadata"]][["Date"]]
    real <- suppressMessages({
      amp_filter_samples(
      combined,
      split_dataset %in% "real",
      Date %in% dates
    )
    })
    combined_test <- amp_merge_ampvis2(
      predicted,
      real,
      by_refseq = FALSE
    )
    combined_test$metadata$predicted <- factor(
      combined_test$metadata$predicted,
      levels = c("real", "predicted")
    )

    #compare real vs predicted (ANOSIM, bray-curtis)
    anosimBC <- vegan::anosim(
      t(combined_test$abund),
      combined_test$metadata$predicted,
      distance = "bray",
      parallel = 30
    )

    #compare real vs predicted (PERMANOVA)
    permanova <- vegan::adonis2(
      t(combined_test$abund) ~ predicted,
      data = combined_test$metadata,
      parallel = 10
    )
    return(
      list(
        anosimBC = anosimBC,
        permanova = permanova
      )
    )
  }
)
names(statistics) <- basename(wwtp_results)
anosimBC_Rstatistics <- sort(
  unlist(
    lapply(
      statistics,
      function(x) x[["anosimBC"]][["statistic"]]
    )
  )
)
anosimBC_Rstatistics

anosimBC_signifs <- sort(
  unlist(
    lapply(
      statistics,
      function(x) x[["anosimBC"]][["signif"]]
    )
  )
)
anosimBC_signifs

permanova <- sort(
  unlist(
    lapply(
      statistics,
      function(x) x[["permanova"]]
    )
  )
)
permanova
```

# Prediction accuracy five number stats for all WWTPs+cluster types+error metrics
```{r fivenum_BC_ASV}
runs <- list.dirs(
  results_batch_dir,
  full.names = TRUE,
  recursive = FALSE
)

d_list <- lapply(runs, read_performance)
names(d_list) <- runs
combined <- rbindlist(
  d_list,
  idcol = "results_folder",
  fill = TRUE
)[
  !is.na(cluster_type) & value > 0
]
#prediction accuracies for each cluster for each dataset+cluster type+error metric
stats <- combined[, .(fivenum_stats = fivenum(value)), by = .(dataset, cluster_type, error_metric)]

# extract medians for BC
mediansBC <- stats[error_metric == "Bray Curtis", .(median = .SD[3,fivenum_stats]), by = c("dataset", "cluster_type")]
mediansBC[, nsamples := as.integer(gsub("[^0-9]", "", dataset))]

# linear relationship between BC+graph accuracy and number of samples
summary(lm(nsamples ~ median, data = mediansBC[cluster_type == "Graph"]))
```
