---
title: "analysis"
author: "KSA"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  error = FALSE,
  fig.width = 12,
  fig.height = 10,
  cache = FALSE
)
require("patchwork")
require("data.table")
require("ampvis2")
require("magrittr")
require("stringr")
require("lubridate")
require("tidygraph")
require("ggraph")
require("igraph")
require("patchwork")
require("scales")

#when knitting will always be done in the context of the Rmd file, otherwise not
if (interactive()) {
  if (!grepl("analysis$", getwd())) {
    setwd("analysis")
  }
}
source("functions.R")
set.seed(42)
# main batch run using run_loopdatasets.bash or run_predwindows.bash
# wrapper scripts. Prediction windows tests are hard coded
results_batchdir <- "results/5pergroup"
```

# five number statistics of sum of reads per data set
```{r fivenum_reads}
datasets <- list.dirs(
  results_batchdir,
  full.names = TRUE,
  recursive = FALSE
)
readstats <- datasets %>%
  lapply(function(dataset) {
    abund <- fread(
      file.path(dataset, "data_reformatted", "abundances.csv"),
      drop = 1
    )
    fivenum(rowSums(abund))
  })
names(readstats) <- basename(datasets)
as.data.frame(readstats, check.names = FALSE)
```

# Prediction accuracy across WWTPs
## Graph model incl graph clustering with 5 per group
```{r}
boxplot_all(
  results_batchdir
)
```

## Check some clusters from AAE
```{r}
plot_performance(
  file.path(results_batchdir, "Aalborg E")
)
```

# Prediction window test
## Aalborg East
```{r}
boxplot_all(
  "results/predwindow_aae",
  add_dataset_info = "predwindow"
)
```

## Randers
```{r}
boxplot_all(
  "results/predwindow_randers",
  add_dataset_info = "predwindow",
  plot_width = 8,
  plot_height = 8
)
```

# PCoA (all WWTPs loop)
```{r PCOA_colored_all}
wwtp_results <- list.dirs(
  results_batchdir,
  full.names = TRUE,
  recursive = FALSE
)

plots <- lapply(
  wwtp_results,
  function(wwtp) {
    plot <- wwtp %>%
      combine_abund(cluster_type = "abund") %>%
      amp_ordinate(
        type = "pcoa",
        distmeasure = "bray",
        transform = "none",
        filter_species = 0,
        sample_color_by = "split_dataset",
        sample_trajectory = "Date",
        sample_trajectory_group = "Date"
      ) +
      scale_color_manual(
        values = c("grey80", RColorBrewer::brewer.pal(3, "Set2")[c(1:3)])
      ) +
      theme(legend.title = element_blank())
    ggsave(
      filename = file.path(
        results_batch_dir,
        paste0("PCoA - ", basename(wwtp), ".png")
      ),
      plot = plot
    )
    return(plot)
  }
)
names(plots) <- basename(wwtp_results)
hideme <- lapply(plots, print)
```

## Time Series examples
```{r timeseries}
wwtp_results <- list.dirs(
  results_batchdir,
  full.names = TRUE,
  recursive = FALSE
)
plots <- lapply(
  wwtp_results,
  function(wwtp) {
    combined <- combine_abund(
      wwtp,
      cluster_type = "abund"
    )
    long <- amp_export_long(
      combined,
      tax_levels = "OTU",
      metadata_vars = c("Date", "split_dataset")
    )

    #check normal distribution
    #qqplot(ASV1, rnorm(1000, mean = mean(ASV1), sd = sd(ASV1)))
    #qqplot(ASV2, rnorm(1000, mean = mean(ASV2), sd = sd(ASV2)))
    ASV1 <- long[grepl("^ASV1;", OTU)]
    if(nrow(ASV1) > 0) {
      plot_asv1 <- plot_timeseries(ASV1, save = FALSE) +
        theme(
          legend.position = "none",
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank()
        ) +
        ggtitle("ASV1 - Tetrasphaera, midas_s_5") +
        scale_y_continuous(breaks = breaks_pretty(n = 5))
    } else {
      plot_asv1 <- plot_spacer()
    }

    ASV2 <- long[grepl("^ASV2;", OTU)]
    if(nrow(ASV2) > 0) {
      plot_asv2 <- plot_timeseries(ASV2, save = FALSE) +
        theme(
          legend.position = "none",
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank()
        ) +
        ggtitle("ASV2 - Ca. Microthrix, parvicella") +
        scale_y_continuous(breaks = breaks_pretty(n = 5))
    } else {
      plot_asv2 <- plot_spacer()
    }

    ASV24 <- long[grepl("^ASV24;", OTU)]
    if(nrow(ASV24) > 0) {
      plot_asv24 <- plot_timeseries(ASV24, save = FALSE) +
        theme(legend.position = "bottom") +
        ggtitle("ASV24 - Nitrospira, defluvii") +
        scale_y_continuous(breaks = breaks_pretty(n = 5))
    } else {
      plot_asv24 <- plot_spacer()
    }

    prediction_example <- plot_asv1 / plot_asv2 / plot_asv24
    ggsave(
      plot = prediction_example,
      filename = file.path(
        results_batch_dir,
        paste0("timeseries example - ", basename(wwtp), ".png")
      ),
      width = 10,
      height = 12
    )

    #plot_timeseries(ASV1, save = TRUE)
    #plot_timeseries(ASV2, save = TRUE)
    #plot_timeseries(ASV3, save = TRUE)
    return(prediction_example)
  }
)
names(plots) <- basename(wwtp_results)
hideme <- lapply(plots, print)

```

# Timeseries examples, prediction window tests
```{r}
results_batch_dir <- "results/predwindow_randers"
runs <- list.dirs(
  results_batch_dir,
  recursive = FALSE,
  full.names = TRUE
)
predwindowtest <- lapply(
  runs,
  function(rundir) {
    d <- amp_export_long(
      combine_abund(
        rundir,
        cluster_type = "abund"
      ),
      tax_levels = "OTU",
      metadata_vars = c("Date", "split_dataset")
    )
    d[, dataset := basename(rundir)]
  }
)
d <- rbindlist(predwindowtest)
d[, dataset := factor(
  dataset,
  levels = unique(dataset)[order(as.integer(gsub("[^0-9]*", "", unique(dataset))))])
]

#ASV1
ASV1 <- d[grepl("^ASV1;", OTU)]
plot_asv1 <- plot_timeseries(
  ASV1,
  save = FALSE
) +
  facet_grid(rows = vars(dataset))
ggsave(
  file = file.path(
    results_batch_dir,
    "predwindowtest_aae_asv1.png"
  ),
  plot = plot_asv1
)

#ASV2
ASV2 <- d[grepl("^ASV2;", OTU)]
plot_asv2 <- plot_timeseries(
  ASV2,
  save = FALSE
) +
  facet_grid(rows = vars(dataset))
ggsave(
  file = file.path(
    results_batch_dir,
    "predwindowtest_aae_asv2.png"
  ),
  plot = plot_asv2
)

#ASV24
ASV24 <- d[grepl("^ASV24;", OTU)]
plot_asv24 <- plot_timeseries(
  ASV24,
  save = FALSE
) +
  facet_grid(rows = vars(dataset))
ggsave(
  file = file.path(
    results_batch_dir,
    "predwindowtest_aae_asv24.png"
  ),
  plot = plot_asv24
)


#plot_asv24 <- plot_timeseries(ASV24, save = FALSE) +
#  theme(legend.position = "bottom") +
#  ggtitle("ASV24 - Nitrospira, defluvii")

```

# Heatmap of real vs predicted
```{r}
results_dir <- "results/5pergroup/Aalborg E"
d <- combine_abund(
  results_dir,
  cluster_type = "graph"
)
test_predicted <- amp_filter_samples(
  d,
  split_dataset %chin% "test",
  predicted %chin% "predicted"
)
test_real <- amp_filter_samples(
  d,
  split_dataset %chin% "real",
  Date %in% test_predicted$metadata$Date
)
test_combined <- amp_merge_ampvis2(
  test_predicted,
  test_real,
  by_refseq = FALSE
)
test_combined$metadata$predicted <- factor(
  test_combined$metadata$predicted,
  levels = c("real", "predicted")
)

realvspredictedheatmap <- amp_heatmap(
  test_combined,
  tax_aggregate = "OTU",
  group_by = c("predicted"),
  normalise = FALSE,
  tax_show = 30,
  plot_values = FALSE,
  facet_by = "Date"
) + theme(strip.text = element_text(angle = 90))
ggsave(
  plot = realvspredictedheatmap,
  filename = file.path(
    results_dir,
    "..",
    paste("realvspredictedheatmap", basename(results_dir), ".png")
  ),
  width = 20,
  height = 10
)
```

# Network plots
```{r}
graph_matrix_dir <- "results/5pergroup/Aalborg E/graph_matrix"
graph_matrix_list <- list.files(
  graph_matrix_dir,
  pattern = "graph_cluster_*",
  full.names = TRUE
)
#list is sorted lexicographically, reorder numerically
order <- graph_matrix_list %>%
  basename %>%
  stringi::stri_extract_all_regex("[0-9]+") %>%
  unlist(use.names = FALSE) %>%
  as.integer %>%
  order
graph_matrix_list <- graph_matrix_list[order]

plot_list <- lapply(
  graph_matrix_list[1:4],
  plot_graph
)

#plots are named after the filenames
plot <- wrap_plots(
  plot_list,
  guides = "collect",
  tag_level = "keep"
)
plot
ggsave(
  plot = plot,
  filename = file.path(
    graph_matrix_dir,
    "..",
    "graph_network_plot.png"
  )
)
```

# Statistical tests of real vs predicted
```{r}
wwtp_results <- list.dirs(
  results_batchdir,
  full.names = TRUE,
  recursive = FALSE
)
statistics <- lapply(
  wwtp_results,
  function(wwtp) {
    #read predicted data (incl train+val)
    combined <- combine_abund(
      wwtp,
      cluster_type = "graph"
    )
    #extract predicted samples (test set only)
    predicted <- amp_filter_samples(
      combined,
      split_dataset %chin% "test",
      predicted %chin% "predicted"
    )
    #extract real samples from the same dates as the predicted
    dates <- predicted[["metadata"]][["Date"]]
    real <- amp_filter_samples(
      combined,
      split_dataset %chin% "real",
      Date %in% dates
    )
    combined_test <- amp_merge_ampvis2(
      predicted,
      real,
      by_refseq = FALSE
    )
    combined_test$metadata$predicted <- factor(
      combined_test$metadata$predicted,
      levels = c("real", "predicted")
    )

    #compare real vs predicted (ANOSIM, bray-curtis)
    anosimBC <- vegan::anosim(
      t(combined_test$abund),
      combined_test$metadata$predicted,
      distance = "bray",
      parallel = 30
    )

    #compare real vs predicted (PERMANOVA)
    permanova <- vegan::adonis2(
      t(combined_test$abund) ~ predicted,
      data = combined_test$metadata,
      parallel = 10
    )
    return(
      list(
        anosimBC = anosimBC,
        permanova = permanova
      )
    )
  }
)
names(statistics) <- basename(wwtp_results)
anosimBC_Rstatistics <- sort(
  unlist(
    lapply(
      statistics,
      function(x) x[["anosimBC"]][["statistic"]]
    )
  )
)
anosimBC_Rstatistics

anosimBC_signifs <- sort(
  unlist(
    lapply(
      statistics,
      function(x) x[["anosimBC"]][["signif"]]
    )
  )
)
anosimBC_signifs

permanova <- sort(
  unlist(
    lapply(
      statistics,
      function(x) x[["permanova"]]
    )
  )
)
permanova
```

# Prediction accuracy five number stats for all WWTPs+cluster types+error metrics
```{r fivenum_BC_ASV}
runs <- list.dirs(
  results_batchdir,
  full.names = TRUE,
  recursive = FALSE
)

d_list <- lapply(runs, read_performance)
names(d_list) <- runs
combined <- rbindlist(
  d_list,
  idcol = "results_folder",
  fill = TRUE
)[
  !is.na(cluster_type) & value > 0
]
#prediction accuracies for each cluster for each dataset+cluster type+error metric
stats <- combined[, .(fivenum_stats = fivenum(value)), by = .(dataset, cluster_type, error_metric)]
stats[error_metric == "Bray Curtis", .(median = .SD[3,fivenum_stats]), by = c("dataset", "cluster_type")]
```

# I dunno
```{r}
combine_results <- function(results_batch_dir) {
  runs <- list.dirs(
      results_batch_dir,
      full.names = TRUE,
      recursive = FALSE
    )
    if (length(runs) == 0) {
      stop("No results folders found, wrong working directory?")
    }

    d_list <- lapply(runs, read_performance)
    names(d_list) <- runs
    combined <- rbindlist(
      d_list,
      idcol = "results_folder",
      fill = TRUE
    )[
      !is.na(cluster_type) & value > 0
    ]
  return(combined)
}

results <- combine_results("results_20220726_divmean/")
results[grepl("Aalborg E", dataset) & cluster_no == 0, ASV := "ASV1"]
results[grepl("Aalborg E", dataset) & cluster_no == 1, ASV := "ASV2"]
results[grepl("Aalborg W", dataset) & cluster_no == 1, ASV := "ASV2"]
results[grepl("Aalborg W", dataset) & cluster_no == 0, ASV := "ASV1"]
results[grepl("Damhusåen-A", dataset) & cluster_no == 1, ASV := "ASV2"]
results[grepl("Damhusåen-A", dataset) & cluster_no == 2, ASV := "ASV1"]
results[grepl("Damhusåen-B", dataset) & cluster_no == 2, ASV := "ASV1"]
results[grepl("Damhusåen-B", dataset) & cluster_no == 0, ASV := "ASV2"]
results[grepl("Damhusåen-C", dataset) & cluster_no == 0, ASV := "ASV2"]
results[grepl("Damhusåen-C", dataset) & cluster_no == 2, ASV := "ASV1"]
results[grepl("Damhusåen-D", dataset) & cluster_no == 2, ASV := "ASV1"]
results[grepl("Damhusåen-D", dataset) & cluster_no == 1, ASV := "ASV2"]
results[grepl("Ejby Mølle", dataset) & cluster_no == 0, ASV := "ASV1"]
results[grepl("Ejby Mølle", dataset) & cluster_no == 8, ASV := "ASV2"]
results[grepl("Mariagerfjord", dataset) & cluster_no == 0, ASV := "ASV1"]
results[grepl("Mariagerfjord", dataset) & cluster_no == 7, ASV := "ASV2"]
results[grepl("Randers", dataset) & cluster_no == 0, ASV := "ASV2"]
results[grepl("Randers", dataset) & cluster_no == 12, ASV := "ASV1"]
results[grepl("Ribe", dataset) & cluster_no == 2, ASV := "ASV1"]
results <- results[ASV %chin% paste0("ASV", 1:2)]

```
